---
title: "HW02"
output: html_document
---
#should we be doing all the assignments on R markdown
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(gapminder)
library(tidyverse)
```
I filtered so that if there was information in Afghanistan, Vietnam, or Cambodia, within the years 1970 and 1979, that information would be filtered into the dataframe named 'gapminder1.1'. I have something loaded that interferes with dplyr, so I have to specify that I am using that. Or is this going && at the end-should i filter for 1970:1979 first? then I would add 

gapminder %>% filter (year == 1970:1979) and then add
gapminder1.1 <- gapminder %>% 
  dplyr::filter(country == 'Afghanistan'| country =='Vietnam' | country == 'Cambodia', year == 1970:1979)
```{r}
gapminder1.1 <- gapminder %>% 
  dplyr::filter(country == 'Afghanistan'| country =='Vietnam' | country == 'Cambodia' && year == 1970:1979)
```
I'm trying to include variables 'country' and gdpPercapita in my filtered version.
```{r}
gapminder1.1 %>%
  dplyr::select(country,gdpPercap)
```
```{r}
gapminder %>%
  dplyr::select(lifeExp)
```
```{r}
gapminderdecrease <- gapminder %>%
  arrange(country,year) %>%
  mutate(laglifeExp = lag(lifeExp,1)) %>%
  mutate (yearlychangelifeExp = lifeExp-laglifeExp) %>%
  filter(yearlychangelifeExp <0) 

increaselifeExp <- gapminder %>%
  arrange(country,year) %>%
  mutate(laglifeExp = lag(lifeExp,1)) %>%
  mutate (yearlychangelifeExp = lifeExp-laglifeExp) %>%
  filter(yearlychangelifeExp >0)

make a new variable that is postitive?, filter that is negative?. dont know what they want different
  
```
filter to decrease, and then add a column saying increase in it?

use mutate to add another column that has the data that I want, which is basically the life expectance column minus the life expectancy of one year before. 
lag life expect gives you the previous year's life expectancy
hopefully this doesnt overlap with the next country

new variable that is the increase in life expectance

top, decrease life expectancy
omcrease lofe expectance (i think that's what they asked for, new variable?)

```{r}
percapmax <- gapminder %>%  
group_by(country) %>% 
      summarise(max=max(gdpPercap))

print(percapmax)
```
this is giving me the highest gdp, but i want the highest gdp PER COUNTRY

Assignment 2

A qualitative and a quantitative value have to be found within gapminder.

A categorical value

```{r}
#i used distinct to determine which continents there were in gapminder
(gapminder %>%
  distinct(continent))

#then i found out how many countries there are in each continent
Continentcount<-(gapminder %>% 
  filter(year==2007) %>%
  count(continent)) 

Continentcount %>% 
  ggplot(aes(x = continent, y = n))+
    geom_bar(stat = "identity") +
    labs(x = "Continent",y = "Number of Countries")
      title = "Life Expectancy vs GDP Per Capita in Canada"
      subtitle = "(based upon data from gapminder in 2007)

```
how many countries, which has the most countries, which has the least countries, which has the median, then bar graph this
geom_
gapminder %>%
mutate(numberofcountries=count(country))
