---
title: "extra"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(gapminder)
library(ggridges)
library(scales)
```
This is task two, where I get the maximum and minimum GDP per capita for all continents. The (dialogue+plot)
I wanna do two, three, and five. try with this one, bar, and next one line.
```{r}
gapminder %>% 
  group_by(continent) %>% 
  summarise(MinGDP=min(gdpPercap),Mincountry=country[gdpPercap==MinGDP],Minyear=year[gdpPercap==MinGDP],MaxGDP=max(gdpPercap),Maxcountry=country[gdpPercap==MaxGDP],Maxyear=year[gdpPercap==MaxGDP])%>% 
  mutate(Range=MaxGDP-MinGDP)

  boxplot(gdpPercap~continent, data=gapminder, notch=TRUE,
  col=(c("gold","darkgreen")),
  main="GDP vs several continents", xlab="Continents", ylab="GDPperCapita")
  
  
```
Task5:
```{r, fig.width = 15, fig.height = 22}

library(gapminder)
library(tidyverse)

gapminder2<-gapminder %>%
  group_by(continent)%>%
  select(continent,gdpPercap,year)%>%
  mutate(laggdpPercap = lag(gdpPercap,1)) %>%
  mutate (yearlygdpPercap = gdpPercap-laggdpPercap)

print(gapminder2)

gapminder2 %>%
  ggplot(aes(yearlygdpPercap, year, color=continent)) +
  geom_line(aes(group = continent), alpha = 0.5) +
  ylab("CO2 Concentration")
 
```

```
task 3
```{r}

```




extra
```{r}
label

ggplot(gapminder, aes((MinGDP=min(gdpPercap)),continent)) +
  ggridges::geom_density_ridges()


ggplot(mpg, aes(cty, hwy)) +
  geom_point() +
  geom_jitter(alpha=0.5,size=1)+
  geom_smooth(method="lm")
  
   gapminder2 %>% 
  filter(continent != "Oceania") %>% 
  ggplot(aes(gdpPercap, lifeExp, size=pop, color=continent)) +
  facet_wrap(~ continent, nrow=1) +
  geom_point(alpha = 0.8) +
  scale_x_log10(labels = scales::comma_format())+
  scale_size_area()

```

