---
title: "extra"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(gapminder)
library(ggridges)
library(scales)
```
This is task two, where I get the maximum and minimum GDP per capita for all continents. The year for minimum GDP was 1952 for Asia, Europe, and Oceania. It was 2002 and 2007 in Africa and the Americas, respectively. The minimum GDP was 10039.59 for Oceania, which is much higher than 241.17 for Africa. The maximum GDP of all of the countries was in the Asia in 2007, with 113523.13 in Kuwait, in 1957. I have made a boplot showing the maximum and minimum GDP per continent. I have also made seperate maximum and minimum graphs which show more detailed numbers, which looks more beautiful, but which I can't combine. I tried to make a column graph that combined the two variables, but I could not figure out how to tell R that I wanted the X axis to be a combination of the minimum and maximum variables.

```{r}
gapminder %>% 
  group_by(continent) %>% 
  summarise(MinGDP=min(gdpPercap),Mincountry=country[gdpPercap==MinGDP],Minyear=year[gdpPercap==MinGDP],MaxGDP=max(gdpPercap),Maxcountry=country[gdpPercap==MaxGDP],Maxyear=year[gdpPercap==MaxGDP])%>% 
  mutate(Range=MaxGDP-MinGDP)

  boxplot(gdpPercap~continent, data=gapminder, notch=TRUE,
  col=(c("gold","darkgreen")),
  main="GDP vs several continents", xlab="Continents", ylab="GDPperCapita")
  
#OR you could look at the max and min seperately, but more closely using the following two graphs:

ggplot(gapminder, aes((MinGDP=min(gdpPercap)),continent,fill=continent)) +
ggridges::geom_density_ridges()+labs(x="Minimum GDP",y="Continent")

ggplot(gapminder, aes((MaxGDP=max(gdpPercap)),continent, fill=continent)) +
ggridges::geom_density_ridges()+labs(x="Maximum GDP",y="Continent")

```
Task5: Here, I showed the life expectancy changes over time by continent. I did this by showing the difference between two years' life expectancy, over many years. I graphed this by coloring based on continent, so you could see a pattern over many years. I changed the figure size to an appropriate height and Width
```{r, fig.width = 10, fig.height = 10}

library(gapminder)
library(tidyverse)

gapminder5<-gapminder %>%
  group_by(continent)%>%
  select(continent,gdpPercap,year)%>%
  mutate(laggdpPercap = lag(gdpPercap,1)) %>%
  mutate (yearlygdpPercap = gdpPercap-laggdpPercap)

print(gapminder5)

gapminder5 %>%
  ggplot(aes(year, yearlygdpPercap, color=continent)) +
  geom_line(aes(group = continent), alpha = 0.5) +
  ylab("GDP per capita")+xlab("Year")
 
```

```
task 3
```{r}

```




extra
```{r}
label




ggplot(mpg, aes(cty, hwy)) +
  geom_point() +
  geom_jitter(alpha=0.5,size=1)+
  geom_smooth(method="lm")
  
   gapminder2 %>% 
  filter(continent != "Oceania") %>% 
  ggplot(aes(gdpPercap, lifeExp, size=pop, color=continent)) +
  facet_wrap(~ continent, nrow=1) +
  geom_point(alpha = 0.8) +
  scale_x_log10(labels = scales::comma_format())+
  scale_size_area()

```

