---
title: "HW04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("tidyverse")
devtools::install_github("tidyverse/tidyr")
library(tidyr)
library(tidyverse)
library(gapminder)
```

count(wt_vars=words) means if there is an option, what 
left is the first seperation, right of dash is the next seperation.- here that is part, dimension, seperated by an underscore.


I am going to pivor wider, and aim to have country on left of dash, and lifeExp on right of dash, per year. 
#now i need to change titles so it says that it is the country's life expectancy( says top tutorial)





I needed to add 3 versions of gapminder, because I needed to plot only the data from above. otherwise, what is y??? because you cant say y is albania and canada? how do you state two y TYPES, itll be something like "part" or dimension" in iris"
is it dbl? double? in iris how does it know what part and dimension it is?

gapminder3 %>% 
    ggplot(aes(x = year, lifeExp)) + 
  geom_point(color = "Blue")+
    labs(
      x="Canada, Albania",
      y="Life Expectancy",
      title = "Life Expectancy per country",
      subtitle = "taken from Gapminder data"
      )
      
      #how do i change color to specific country
gapminder2 %>% 
  ggplot(aes(x = year, y = lifeExp)) + 
  geom_point(color = Canada=="blue", color=Albania=="green")+
    labs(
      x="Canada, Albania",
      y="Life Expectancy",
      title = "Life Expectancy per country",
      subtitle = "taken from Gapminder data"
      )      
      
#scatterplot-thats geompoint
#relengthen

#do i need the print??

```{r}
print(
gapminderlife<-gapminder%>%
  select(year,country,lifeExp)%>%
  dplyr::filter(country=="Canada"|country=="Albania")%>% 
  pivot_wider(id_cols     = c(-country,-lifeExp),
              names_from  = country, 
              values_from = lifeExp)
)


gapminderlife %>% 
  ggplot(aes(x = Canada, y = Albania)) + 
  geom_point() + # 
    labs(
      x="Canada Life Expectancy in years",
      y="Albania Life Expectancy in years",
      title = "Life Expectancy of Canada vs Albania",
      subtitle = "taken from Gapminder data"
      ) 


print
(gapminderlifelong<-gapminderlife %>% 
  pivot_longer(cols     = (-year),
              names_to  = "country", 
              values_to = "lifeExp") 
)
  

```
ASSIGNMENT 2

```{r}
gapmindermulti<-gapminder%>%
  dplyr::filter(country=="Canada"|country=="Albania")%>% 
  pivot_wider(id_cols     = c(country,year),
              names_from  = country, 
              names_sep   = "_",
              values_from = c(lifeExp,gdpPercap))

gapmindermulti

#expand

gapmindermultilong <-gapmindermulti %>% 
  pivot_longer(cols     = (-year),
              names_sep = "_", 
              names_to = c(".value", "country"))

gapmindermultilong  

```
```{r}



```
```{r}
#assignment 3

guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")

```
```{r}

#Extra

gapminder2 %>% 
  ggplot(aes(x = year, y = lifeExp, colour = country)) + 
  geom_point()+
    labs(
      x="Year",
      y="Life Expectancy",
      title = "Life Expectancy per country",
      subtitle = "taken from Gapminder data"
      )    
```

