---
title: "HW04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
install.packages("tidyverse")
devtools::install_github("tidyverse/tidyr")
library(tidyr)
library(tidyverse)
library(gapminder)
```

count(wt_vars=words) means if there is an option, what 
left is the first seperation, right of dash is the next seperation.- here that is part, dimension, seperated by an underscore.


I am going to pivor wider, and aim to have country on left of dash, and lifeExp on right of dash, per year. 
#now i need to change titles so it says that it is the country's life expectancy( says top tutorial)





I needed to add 3 versions of gapminder, because I needed to plot only the data from above. otherwise, what is y??? because you cant say y is albania and canada? how do you state two y TYPES, itll be something like "part" or dimension" in iris"
is it dbl? double? in iris how does it know what part and dimension it is?

gapminder3 %>% 
    ggplot(aes(x = year, lifeExp)) + 
  geom_point(color = "Blue")+
    labs(
      x="Canada, Albania",
      y="Life Expectancy",
      title = "Life Expectancy per country",
      subtitle = "taken from Gapminder data"
      )
      
      #how do i change color to specific country
gapminder2 %>% 
  ggplot(aes(x = year, y = lifeExp)) + 
  geom_point(color = Canada=="blue", color=Albania=="green")+
    labs(
      x="Canada, Albania",
      y="Life Expectancy",
      title = "Life Expectancy per country",
      subtitle = "taken from Gapminder data"
      )      
      
#scatterplot-thats geompoint
#relengthen

```{r}
gapminder2<-gapminder%>%
  select(year,country,lifeExp)%>%
  dplyr::filter(country=="Canada"|country=="Albania")

gapminder2

gapminder3<-gapminder2%>% 
  pivot_wider(id_cols     = c(-country,-lifeExp),
              names_from  = country, 
              values_from = lifeExp)


gapminder3

gapminder2 %>% 
  ggplot(aes(x = year, y = lifeExp, colour = country)) + 
  geom_point()+
    labs(
      x="Year",
      y="Life Expectancy",
      title = "Life Expectancy per country",
      subtitle = "taken from Gapminder data"
      )    

#relengthen

gapminder3 %>% 
  pivot_wider(id_cols     = c(year),
              names_from  = c(country, lifeExp), 
              names_sep   = "_", 
              values_from = country)




```
ASSIGNMENT 2

```{r}
#country isnt a thing, so what are y values called????


(guest_long <- guest %>% 
  pivot_longer(cols      = c(-party,-name), 
               names_to  = c(".value","event"),
               names_sep = "_"))

guest_long %>% 
  pivot_wider(id_cols     = c(party, name), 
              names_from  = event, 
              names_sep   = "_", 
              values_from = c(meal, attendance))


#can you have 1952 canada and 1952 albania

gg %>% 
  ggplot(aes(x ))
```
```{r}
gapmindermulti<-gapminder%>%
  select(year,country,lifeExp,gdpPercap)%>%
  dplyr::filter(country=="Canada"|country=="Albania")

gapmindermulti

gapmindermulti2<-gapmindermulti%>% 
  pivot_wider(id_cols     = c(years),
              names_from  = country, 
              values_from = lifeExp,gdpPercap)


```
```{r}

```
```{r}
gapminder_wide <- gapminder %>% 
  filter(country %in% c("Canada", "China")) %>% # Filter for rows with Canada or China
  pivot_wider(id_cols = c(country, year), # Keep countries and year columns
              values_from = lifeExp, # Cells of new columns will contain lifeExp values
              names_from = country) # Newly generated columns are named by the values
                                    # of the country column

gapminder_wide %>% 
  ggplot(aes(x = Canada, y = China)) + # Specify the x and y variables for ggplot
  geom_point() + # Add point aesthetics for each data point
  xlab("Canada life expectancy (yrs)") + # label x axis
  ylab("China life expectancy (yrs)") + # label y axis
  ggtitle("Correlation between the life expectancies of China and Canada")
```

